\documentclass[12pt,a4paper]{article}

% Packages
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{tcolorbox}
\usepackage{fancyhdr}
\setlength{\headheight}{15pt}


% Header and footer
\pagestyle{fancy}
\fancyhf{}
\rhead{Mathematics for Computer Science }
\lhead{Vedant Patil}
\cfoot{\thepage}

% Title
\title{Lecture Notes: Induction Proofs}
\author{Vedant Patil}
\date{\today}

\begin{document}

\maketitle

\section{Overview}
\begin{tcolorbox}[colback=yellow!10!white,colframe=yellow!50!black,title=Key Points]
  \begin{itemize}
    \item Key point 1
    \item Key point 2
    \item Key point 3
  \end{itemize}
\end{tcolorbox}

\section{Detailed Notes}
\subsection{Proof by Condtradiction}
To prove that a proposition P is true, we assume that P is false, formally determined as Not P is True. then use that hypothesis to derive a falsehood or condtradiction. If \( \not \) P \( \Rightarrow \) F is true, the only way for this state ment to be true is if not P is false or otherwise if P is true.

\subsection{Subtopic 2}
Your notes for subtopic 2

\section{Important Formulas/Theorems/Definitions}
\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=Key Formula/Theorem]
  State an important formula or theorem here
\end{tcolorbox}

\section{Examples}
\begin{tcolorbox}
\textbf{Proof that $\sqrt{2}$ is irrational}

Assume, for the purpose of contradiction, that $\sqrt{2}$ is rational.

\begin{enumerate}
    \item If $\sqrt{2}$ is rational, then $\sqrt{2} = \frac{a}{b}$ for some integers $a$ and $b \neq 0$, with $\frac{a}{b}$ in lowest terms.
    \item Squaring both sides: $2 = \frac{a^2}{b^2}$
    \item Multiply both sides by $b^2$: $2b^2 = a^2$
    \item Therefore, $a^2$ is even, which implies $a$ is even.
    \item Let $a = 2k$ for some integer $k$.
    \item Substituting: $2b^2 = (2k)^2 = 4k^2$
    \item Simplifying: $b^2 = 2k^2$
    \item Therefore, $b^2$ is even, which implies $b$ is even.
    \item But if both $a$ and $b$ are even, the fraction $\frac{a}{b}$ is not in lowest terms.
    \item This contradicts our initial assumption.
    \item \( \rightthreetimes \)
\end{enumerate}

Therefore, $\sqrt{2}$ cannot be expressed as a ratio of integers and is thus irrational.
\end{tcolorbox}

\section{Questions/Topics for Further Study}
\begin{itemize}
  \item Question or topic for further study
\end{itemize}

\end{document}
